Qubit tests from classical codes

### 2022-12-19 quantum_codes_test.tex
clarified connection with dlS game

Let $\code$ be an $[n,k,d]$ linear code and $M$ an $r$-local tester for $\code$. 
We use $G_{\dlS}$ to denote the game introduced in~\cite[Section 3.4]{de2022spectral}, specialized to the following choices. The group is $H=\field^k$. Let
\[ S_X=S_Z=\{G_\code e_i:\,i\in\{1,\ldots,n\}\}\subseteq \field^k\;,\]
and let $\mu_{\dlS}$ be the uniform distribution over $S_X\times S_Z$. Let $\Omega$ be the support of $\mu_{\dlS}$ and $\alpha,\beta$ the coordinate projections. Then $G_{\dlS}=(\mX,\mu_{\dlS},\mA,D)$ has question set $\mX = \{PX,PZ\} \cup (\mX_{com}\times \Omega_+) \cup (\mX_{anticom} \times \Omega_-)$ and is as described in~\cite[Section 3.4]{de2022spectral}.

@@

\begin{figure}[!htbp]
  \centering
  \begin{gamespec}
Let $M$ be an $r$-local tester for the $[n,k,d]$ code $\code$. 
    \begin{enumerate}
      \setlength\itemsep{1pt}
    \item (\textbf{Code test}) Select $W\in \{X,Z\}$ uniformly at random, and $\rand$ a setting for $M$'s random tape at random. Let $S=M(\rand)\subseteq\{1,\ldots,n\}$. Sample $i\in S$ uniformly at random. Send $(W,S)$ to $\alice$ and $(W,i)$ to $\bob$. Receive $a\in \field^S$ from $\alice$ and $b\in \field$ from $\bob$. Accept if and only if $M(S,a)=1$ and $a_i = b$.  
    \item (\textbf{Anti-commutation test}) Select $(i_X,i_Z)\in \{1,\ldots,n\}^2 $ uniformly at random. Let $\omega = (G_\code e_{i_X}, G_\code e_{i_Z})$ and $\gamma =  (G_\code e_{i_X}) \cdot(G_\code e_{i_Z}) \in \field$.
		\begin{enumerate}
		\item If $\gamma=0$ then select a pair of questions $(x_c,y_c)$ as in the commutation game. Send $(x_c,\omega)$ to $\alice$ and $(y_c,\omega)$ to Bob. Accept if and only if their answers are accepted in the commutation game. 
		\item If $\gamma\neq 0$ then do the same but for the anti-commutation game. 
		\end{enumerate} 
		 \item (\textbf{Consistency test}) Select $(i_X,i_Z)\in \{1,\ldots,n\}^2 $ and $W\in \{X,Z\}$ uniformly at random. Let $\omega=(G_\code e_{i_X}, G_\code e_{i_Z})$ and $\gamma = (G_\code e_{i_X}) \cdot(G_\code e_{i_Z}) \in \field$. Send $(W,i_W)$ to $\alice$ and $(x_{W,\gamma},\omega)$ to $\bob$. Receive $a\in \field$ and $b\in \field$ respectively. Accept if and only if $a=b$. 
    \end{enumerate}
  \end{gamespec}
  \caption{The braiding test over $\code$ verifies that the players respond consistently with a uniformly random codeword from $\code$, obtained by measuring $k$ shared EPR pairs in the standard or Hadamard basis and encoding the resulting $k$-bit string.}
  \label{fig:braiding-test}
\end{figure}

@@

\underline{Soundness:} Next we show soundness. Let $\strategy$ be a synchronous strategy for $G$  in  $(\mM,\tau)$ that succeeds with probability at least $1-\eps$. For $W\in\{X,Z\}$ let $\strategy^W$ be the strategy in $G_{\code,M}$ that is obtained by restricting $\strategy$ to the relevant measurements, i.e.\ the $P^{W,S}$ and $P^{W,i}$.  Then $\strategy^W$ succeeds with probability at least $1-3\eps$ in $G_{\code,M}$. Using the assumption that $M$ is $\delta(\cdot)$-quantum sound and the first item from Proposition~\ref{prop:sound-game} it follows that there is a $\delta_1 = O(r\sqrt{\delta(6\eps)})$ such that for each $W\in\{X,Z\}$, $\strategy^W$ is $(\delta_1,\mu)$ close to a perfect strategy $\tilde{\strategy}^W$ on $(\mN^W,\tau^W)$. This strategy in particular has measurement operators $\{ \tilde{P}^{W,S}_a\}$ and $\{\tilde{P}^{W,i}_b\}$ associated with questions of the form $S$ and $i$ in $G_{\code,M}$. Since $(\code,M)$ is Abelian, there is a  POVM $\{\tilde{P}^W_u\}_{u\in \field^n}$ such that $\sum_{u\in \code}\tilde{P}^W_u = \Id$ and for each $i\in \{1,\ldots,n\}$, $\tilde{P}^{W,i}_b = \sum_{a:a_i=b} \tilde{P}^W_a$.% where  $u_i = G_\code e_i$.

Applying Lemma~\ref{lem:pull-back}, we obtain projective measurements $\{Q^W_u\}$ on $\mM$ such that 
\begin{equation*}
\sum_u \big\|Q^W_u - (w^W)^* \tilde{P}_u^W w^W \big\|^2_2 \,=\, O(\delta_1)\;.
\end{equation*}
For any $i\in\{1,\ldots,n\}$ let $Q^{W,i}_b = \sum_{a:\,a_i=b}  Q^W_a$. Then by Lemma~\ref{lem:dp},
\begin{align*}
\Es{i}\sum_b \tau\big( Q^{W,i}_b (w^W)^*\tilde{P}^{W,i}_b(w^W) \big)
&\geq \Es{i}\sum_a \tau\big( Q^{W}_a (w^W)^*\tilde{P}^{W}_a(w^W) \big)\\
&\geq 1-O(\delta_1)\;.
\end{align*}
For $W\in\{X,Z\}$ and $b\in \F^k$ let $R^W_b = Q^W_{G_\mC^T b}$, where by definition $G_\mC^T b\in \mC$. 
Applying Lemma~\ref{lem:close-value} we deduce that the strategy that consists of $R^X$, $R^Z$, $P^{x_c,\omega}$ for $(x_c,\omega) \in \mX_{com}\times \Omega_+$ and $P^{x_{ac},\omega}$ for $(x_{ac},\omega)\in \mX_{anticom} \times \Omega_-$ succeeds in the game $G_{\dlS}$ with probability $1-O(\sqrt{\delta_1})$\tnote{should justify this with more detail?}. Applying~\cite[Corollary 3.9]{de2022spectral}, the conclusion follows. 
\end{proof}

>>>>

Let $\code$ be an $[n,k,d]$ linear code and $M$ an $r$-local tester for $\code$. 
We use $G_{\dlS}$ to denote the game introduced in~\cite[Section 3.4]{de2022spectral}, specialized to the following choices. The group is $H=\field^k$. Let
\[ S_X=S_Z=\{G_\code e_i:\,i\in\{1,\ldots,n\}\}\subseteq \field^k\;,\]
and let $\mu_{\dlS}$ be the uniform distribution over $S_X\times S_Z$. Let $\Omega$ be the support of $\mu_{\dlS}$ and $\alpha,\beta$ the coordinate projections. Then $G_{\dlS}=(\mX,\mu_{\dlS},\mA,D)$ has question set $\mX = \{PX,PZ\} \cup (\mX_{com}\times \Omega_+) \cup (\mX_{anticom} \times \Omega_-)$ and is as described in~\cite[Section 3.4]{de2022spectral}. For clarity we recall the game, using our notation, in Figure~\ref{fig:dlS}.

\begin{figure}[!htbp]
  \centering
  \begin{gamespec}
Let $S_X,S_Z\subseteq \field^k$.  Select $\omega = (\omega_X,\omega_Z)\in S_X \times S_Z $ uniformly at random. Let $\gamma = \omega_X \cdot \omega_Z \in \field$. Execute either of the following tests with probability $1/3$ each. 
    \begin{enumerate}
      \setlength\itemsep{1pt}
    \item (\textbf{Anti-commutation test}) 
		\begin{enumerate}
		\item If $\gamma=0$ then select a pair of questions $(x_c,y_c)$ as in the commutation game. Send $(x_c,\omega)$ to $\alice$ and $(y_c,\omega)$ to Bob. Accept if and only if their answers are accepted in the commutation game. 
		\item If $\gamma\neq 0$ then do the same but for the anti-commutation game. 
		\end{enumerate} 
		 \item (\textbf{$Z$-Consistency test}) Send $Z$ to $\alice$ and $(x_{Z,\gamma},\omega)$ to $\bob$. Receive $a\in \field^k$ and $b\in \field$ respectively. Accept if and only if $a\cdot \omega_Z=b$. 
		 \item (\textbf{$X$-Consistency test}) Send $X$ to $\alice$ and $(x_{X,\gamma},\omega)$ to $\bob$. Receive $a\in \field^k$ and $b\in \field$ respectively. Accept if and only if $a\cdot \omega_X=b$. 
    \end{enumerate}
  \end{gamespec}
  \caption{The game $G_{\dlS}$ checks (anti)commutation relations between two collections of observables.}
  \label{fig:dlS}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering
  \begin{gamespec}
Let $M$ be an $r$-local tester for the $[n,k,d]$ code $\code$.  Execute either of the following tests with probability $1/3$ each. 
    \begin{enumerate}
      \setlength\itemsep{1pt}
    \item (\textbf{Code test}) Select $W\in \{X,Z\}$ uniformly at random, and $\rand$ a setting for $M$'s random tape at random. Let $S=M(\rand)\subseteq\{1,\ldots,n\}$. Sample $i\in S$ uniformly at random. Send $(W,S)$ to $\alice$ and $(W,i)$ to $\bob$. Receive $a\in \field^S$ from $\alice$ and $b\in \field$ from $\bob$. Accept if and only if $M(S,a)=1$ and $a_i = b$.  
    \item (\textbf{Anti-commutation test}) Select $(i_X,i_Z)\in \{1,\ldots,n\}^2 $ uniformly at random. Let $\omega = (G_\code e_{i_X}, G_\code e_{i_Z})$ and $\gamma =  (G_\code e_{i_X}) \cdot(G_\code e_{i_Z}) \in \field$.
		\begin{enumerate}
		\item If $\gamma=0$ then select a pair of questions $(x_c,y_c)$ as in the commutation game. Send $(x_c,\omega)$ to $\alice$ and $(y_c,\omega)$ to Bob. Accept if and only if their answers are accepted in the commutation game. 
		\item If $\gamma\neq 0$ then do the same but for the anti-commutation game. 
		\end{enumerate} 
		 \item (\textbf{Consistency test}) Select $(i_X,i_Z)\in \{1,\ldots,n\}^2 $ and $W\in \{X,Z\}$ uniformly at random. Let $\omega=(G_\code e_{i_X}, G_\code e_{i_Z})$ and $\gamma = (G_\code e_{i_X}) \cdot(G_\code e_{i_Z}) \in \field$. Send $(W,i_W)$ to $\alice$ and $(x_{W,\gamma},\omega)$ to $\bob$. Receive $a\in \field$ and $b\in \field$ respectively. Accept if and only if $a=b$. 
    \end{enumerate}
  \end{gamespec}
  \caption{The braiding test over $\code$ verifies that the players respond consistently with a uniformly random codeword from $\code$, obtained by measuring $k$ shared EPR pairs in the standard or Hadamard basis and encoding the resulting $k$-bit string.}
  \label{fig:braiding-test}
\end{figure}

@@

\underline{Soundness:} Next we show soundness. Let $\strategy$ be a synchronous strategy for $G$  in  $(\mM,\tau)$ that succeeds with probability at least $1-\eps$. For $W\in\{X,Z\}$ let $\strategy^W$ be the strategy in $G_{\code,M}$ that is obtained by restricting $\strategy$ to the relevant measurements, i.e.\ the $P^{W,S}$ and $P^{W,i}$.  Then $\strategy^W$ succeeds with probability at least $1-3\eps$ in $G_{\code,M}$. Using the assumption that $M$ is $\delta(\cdot)$-quantum sound and the first item from Proposition~\ref{prop:sound-game} it follows that there is a $\delta_1 = O(r\sqrt{\delta(6\eps)})$ such that for each $W\in\{X,Z\}$, $\strategy^W$ is $(\delta_1,\mu)$ close to a perfect strategy $\tilde{\strategy}^W$ on $(\mN^W,\tau^W)$. This strategy in particular has measurement operators $\{ \tilde{P}^{W,S}_a\}$ and $\{\tilde{P}^{W,i}_b\}$ associated with questions of the form $S$ and $i$ in $G_{\code,M}$. These satisfy in particular that 
\begin{equation}\label{eq:main-0}
\Es{i\in\{1,\ldots,n\}}\sum_b \big\|P^{W,i}_b - (w^W)^* \tilde{P}_i^W w^W \big\|^2_2 \,=\, O(\delta_1)\;.
\end{equation}
Furthermore, since $(\code,M)$ is Abelian, there is a  POVM $\{\tilde{P}^W_u\}_{u\in \field^n}$ such that $\sum_{u\in \code}\tilde{P}^W_u = \Id$ and for each $i\in \{1,\ldots,n\}$, $\tilde{P}^{W,i}_b = \sum_{a:a_i=b} \tilde{P}^W_a$.% where  $u_i = G_\code e_i$.

Applying Lemma~\ref{lem:pull-back}, we obtain projective measurements $\{Q^W_u\}$ on $\mM$ such that 
\begin{equation}\label{eq:main-1}
\sum_u \big\|Q^W_u - (w^W)^* \tilde{P}_u^W w^W \big\|^2_2 \,=\, O(\delta_1)\;.
\end{equation}
For any $i\in\{1,\ldots,n\}$ let $Q^{W,i}_b = \sum_{a:\,a_i=b}  Q^W_a$. Then by Lemma~\ref{lem:dp},
\begin{align*}
\Es{i}\sum_b \tau\big( Q^{W,i}_b (w^W)^*\tilde{P}^{W,i}_b(w^W) \big)
&\geq \Es{i}\sum_a \tau\big( Q^{W}_a (w^W)^*\tilde{P}^{W}_a(w^W) \big)\\
&\geq 1-O(\delta_1)\;.
\end{align*}
For $W\in\{X,Z\}$ and $b\in \F^k$ let $R^W_b = Q^W_{G_\mC^T b}$, where by definition $G_\mC^T b\in \mC$.

We now define a strategy $\strategy'$ for the game $G_{\dlS}$. On question $W\in \{X,Z\}$ the projective measurement is $\{R^W_b\}$. On question of the form $(x_c,\omega)$ for $x_c$ a question in the commutation game, the projective measurement is $\{P^{x_c,\omega}\}$, i.e.\ the same projective measurement as used in $\strategy$. Similarly, on a question of the form $(x_{ac},\omega)$ for $x_{ac}$ a question in the anti-commutation game, the projective measurement is $\{P^{x_{ac},\omega}\}$.

To conclude we show that this strategy succeeds in the game $G_{\dlS}$ with probability $1-O(\sqrt{\delta_1})$. Assuming that this has been shown, by~\cite[Corollary 3.9]{de2022spectral} the strategy $\strategy'$ is $O(\sqrt{\delta_1})$-close to a strategy $\strategy''$ on an algebra of the form $(M_{2^{k}}(\C)\otimes \mN,\tr\otimes \tau')$ such that $P^W_b = \tau^W_b\otimes 1_\mN$. By definition of $R$, 
\begin{equation}\label{eq:main-3}
 Q^{W,i}_b \,=\,  \sum_{a:\,a_i=b}  Q^W_a \,=\, \sum_{c: (G_\mC^T c)_i=b}  R^W_c \;,
\end{equation}
hence
\begin{equation*}
 \widehat{Q^{W,i}}\,=\, \sum_c (-1)^{c\cdot (G_\mC e_i)} R^W_c \,=\, \widehat{R^W}(G_\mC e_i)\;.
\end{equation*}
Using the definition of the game distribution, closeness of $\strategy'$ and $\strategy''$ thus implies that
\begin{equation*}
\Es{i\in\{1,\ldots,n\}} \big\|\widehat{Q^{W,i}} - (w'')^* {\tau^W}(G_\mC e_i) (w'') \big\|_2^2 \,=\,O(\sqrt{\delta_1})\;.
\end{equation*}
Combining with~\eqref{eq:main-0} and~\eqref{eq:main-1}, this shows the theorem.

It remains to verify that $\strategy'$ succeeds in the game $G_{\dlS}$ with probability $1-O(\sqrt{\delta_1})$. By definition $\strategy'$ succeeds in the (anti)-commutation test with probability $1-O(\eps)$. It remains to check the $W$-consistency test, for $W\in\{X,Z\}$. Because $\strategy$ succeeds with probability $1-O(\eps)$ in the consistency test, 
\begin{equation*}
\Es{i_X,i_Z\in\{1,\ldots,n\}} \sum_b \tau\big( P^{W,i}_b P^{x_{W,\gamma},\omega}_b\big) \,\geq\, 1-O(\eps)\;,\
\end{equation*}
where $\omega$ and $\gamma$ are defined as in Figure~\ref{fig:braiding-test}. Using~\eqref{eq:main-0},~\eqref{eq:main-1}
and Lemma~\ref{lem:close-value} it follows that 
\begin{equation*}
\Es{i_X,i_Z\in\{1,\ldots,n\}} \sum_b \tau\big( Q^{W,i}_b P^{x_{W,\gamma},\omega}_b\big) \,\geq\, 1-O(\sqrt{\delta_1})\;,\
\end{equation*}
Using~\eqref{eq:main-3}, this can be rewritten as 
\begin{equation}\label{eq:main-4}
\Es{i_X,i_Z\in\{1,\ldots,n\}} \sum_{b,c: (G_\mC^T c)_i=b} \tau\big( R^{W}_c P^{x_{W,\gamma},\omega}_b\big) \,\geq\, 1-O(\sqrt{\delta_1})\;.
\end{equation}
Since $\omega_W = G_\mC e_i$, $(G_\mC^T c)_i = c\cdot \omega_W$. Thus~\eqref{eq:main-4} shows that $\strategy'$ succeeds with probability $1-O(\sqrt{\delta_1})$ in the $W$-consistency test, as desired. 
\end{proof}
